FinalManuscript_DISCS_UG

### 2018-10-07 parts/02_relatedliterature_03facialrecognition.tex

\subsection{Haar Cascades [WIP]}
A \textit{Haar cascade classifier} is another method for facial recognition by Paul Viola and Michael Jones. It uses Haar-like features to detect regularities in a face image. A \textit{Haar-like feature} looks at a certain rectangular region of the image through a filter-like window.

@@

A Haar-like feature is composed of black and white rectangles, and arranged such that it detects a presence or an absence of a particular image feature or regularity of interest. For example, a feature may detect the difference in intensities between the eyes and the cheeks, another feature would detect the nose bridge, and another one would detect depending on the location of the eyes. The detection is done by summing up the intensities within the black region and the white region, and then calculating the difference.

@@

Most of the Haar-like features are irrelevant, because a feature that works in a certain part of the image would not necessarily work in another part. The whole process may become needlessly inefficient due to computational time being potentially wasted on a huge number of irrelevant features, even though computing for a feature is said to be efficient. That is why we need to select the best features and use those to form a strong classifier in a process called AdaBoost.

In itself, a Haar feature is considered as a \textit{weak classifier}, where the probability of detecting a certain feature is slightly greater than detecting at random. Because of this, a large number of Haar features must be used together by forming them into cascade classifiers in order to achieve a sufficient enough accuracy. The resulting strong classifier would then be a linear combination of the weak classifiers.

\[h \left(x, f, p, \theta \right) = \begin{cases}1 & \text{if } pf(x) < p\theta\\ 0 & \text{otherwise}\end{cases}
\]

>>>>

\subsection{Haar Cascades}
A \textit{Haar cascade classifier} is another method for facial recognition by Paul Viola and Michael Jones \cite{viola_robust_2004}. It uses Haar-like features to detect regularities in a face image. A \textit{Haar-like feature} looks at a certain rectangular region of the image through a filter-like window.

@@

A Haar-like feature is composed of black and white rectangles, and arranged such that it detects a presence or an absence of a particular image feature or regularity of interest. For example, a feature may detect the difference in intensities between the eyes and the cheeks, another feature would detect the edges of the face, and another one would detect depending on the location of the eyes. The detection is done by summing up the intensities within the black region and the white region, and then calculating the difference.

@@

Most of the Haar-like features are irrelevant, because a feature that works in a certain part of the image would not necessarily work in another part. The whole process may become needlessly inefficient due to computational time being potentially wasted on a huge number of irrelevant features, even though computing for a feature is said to be efficient. That is why we need to select the best features and use those to form a strong classifier in a process called AdaBoost \cite{viola_robust_2004}.

In itself, a Haar-like feature is considered as a \textit{weak classifier}, the probability of detecting a certain feature is slightly greater than detecting at random \cite{viola_robust_2004}. So a large number of Haar-like features must be used together by forming them into cascade classifiers in order to achieve a sufficient enough accuracy. The resulting strong classifier would then be a linear combination of the weak classifiers.

Each stage in the cascade classifier consists of weak classifiers that uses decision stumps, so that the weak classifiers can classify many examples correctly. A \textit{decision stump} can be defined as
\[h \left(x, f, p, \theta \right) = \begin{cases}1 & \text{if } pf(x) < p\theta\\ 0 & \text{otherwise}\end{cases}
\]
where $x$ is a fixed window of the image, $f$ is the feature, $p$ is the polarity, and $\theta$ is the threshold \cite{viola_robust_2004}.

Each stage also labels the region as positive (meaning that there is a face within the region), and negative (if none). If the detector sees the region as negative, it is immediately discarded and not processed anymore. Otherwise if the region is labeled as positive, it moves to the next stage for further processing. When a region successfully completes all the stages, it can be considered as a face region.
